The complex structure and peculiar internal motions of the planetary nebula NGC 6210

### 2020-03-24 ms.tex
Updates from Overleaf
Compare proper motions with Schonberner:2018a

\usepackage{xcolor}
\usepackage{hyperref}
\usepackage{siunitx}
\usepackage{newtxtext}
\usepackage[stix2,smallerops]{newtxmath}
\usepackage{booktabs}
\hypersetup{colorlinks=True, linkcolor=blue!50!black, citecolor=black,
  urlcolor=blue!50!black}
\usepackage{etoolbox}
\robustify\bfseries
\robustify\itshape

@@

%\author{Ma.\ T. Garc\'{\i}a-D\'{\i}az, J. A. L\'opez, W. Steffen \& 
%  M. G., Richer 
% \affil{Instituto de Astronom\'ia,
%    Universidad Nacional Aut\'onoma de M\'exico} Campus Ensenada,
%  Ensenada, Baja California, 22800, M\'exico}
%\affil{tere@astrosen.unam.mx, jal@astrosen.unam.mx, wsteffen@astrosen.unam.mx, richer@astrosen.unam.mx}

@@

\begin{abstract}
  A comprehensive kinematic study has been carried out on the planetary nebula (PN) NGC 6210. Multiple long-slits, echelle spectra have been obtained over the face of this nebula mapping its full shell structure, the opposite pairs of extended, twisted arms and the bipolar collimated outflows and bullets. Public {\it HST} imagery has been used to identify kinematic elements with structural components. The long-slit spectroscopic information has been combined into channel maps that greatly facilitate visualizing the otherwise intricate expansion pattern of this  planetary nebula. The global morphology of NGC 6210 is  reminiscent of other planetary nebulae with extended X-ray emission, suggesting the presence of a central hot bubble produced by shocked stellar wind. In spite of the dramatic structure of this PN substantial expanding radial motions are only found in the material surrounding the central, inner shell. The slow radial motions of the collimated outflows and the bullet-like knots indicate that they are  moving away from the core very close to the plane of the sky, nearly perpendicular to the line of sight
\end{abstract}

@@

\section{INTRODUCTION}
\label{sec:introduction}
NGC 6210 is a bright and relatively large PN in the northern sky. The first photographic image of NGC 6210 was published by \citet{Duncan:1937a}. In that old, fuzzy image a pair of twisted arms protruding in opposite directions from a blurred,  bright, nebular core are apparent. Several studies have discussed the inner kinematic motions of this object, e.g. \citet{Osterbrock:1966a, Weedman:1968a, Becker:1984a, Icke:1989a}. However, none of those studies had the spatial coverage and spectral resolution needed to produce a reliable spatio-kinematic model of this PN. 
The best kinematic study to date on NGC 6210 is probably the one published by \citet{Phillips:1996a}  who obtained 11 long-slits centered on the nebular core, with each slit at different position angles thus providing a good spatial coverage but with only limited, low and medium spectral resolution. From these data  they suggested a  reasonable kinematic model for the ionized nebula. NGC 6210 jumped into fame thanks to the {\it HST} images obtained during the 1996 - 1998 period under the observing programs 6347 (K. Borkowski), 6792 (R. Rubin), 7501 (A. Hajian)  and 11122 (B. Balick). These images showed for the first time the extraordinarily complex morphology of NGC 6210. Incidentally the  {\it HST} and ground-based imagery prompted the nickname "the turttle" for this nebula, given the shape of its main body and extended arms that resemble the shell of a turttle with a long neck and its fins. The distance estimates to NGC 6210 vary roughly from 1.5 to 2.0 kpc. \citep{Hajian:1995a} derive a distance to NGC 6210 from VLA 6 cm  expansion parallax measurements of 1.57$\pm0.40$ kpc, considering an expansion distance of 23 \kms

In this work we present a thorough spectral mapping of all the morphological elements of the PN obtained at high spectral resolution that allow 
disentangling the various spatio-kinematic components and provide and overall view of its structure and evolution. Radial velocities are combined with 2-epoch HST \oiii and \nii archive image sets (http://archive.stsci.edu/) to derive proper motions. Individual long-slit echelle data obtained over N observing runs are combined into emission and velocity channel maps that serve as an excellent tool to visualize and understand the peculiar velocity fields, of the several structural components.

In Section 2, we describe our observations and the data-reduction steps, In Section 3 we describe the velocity-channel map of emission. In Section 4, we discuss the radial velocities and proper motions of the nebula. In Seccion 5 we present the discussion and conclusion.

@@

Long-slit, echelle, spectroscopic observations of the nebula NGC~6210
were performed with the 2.1~m telescope at the Observatorio
Astron\'omico Nacional at San Pedro M\'artir, (OAN-SPM), Baja
California, M\'exico. We used the Manchester Echelle Spectrometer
(MES-SPM) \citep{Meaburn:2003a} on the 2.1 m telescope in its $f$/7.5
configuration.  The MES-SPM is a long-slit, echelle spectrometer that
has no cross-disperser; it isolates single orders using interference
filters. For the 2003, 2004 and 2011 of the present observations, we used a
SITE-3 CCD detector with 1024 $\times$ 1024 square pixels, each 24
$\mu$m on a side ($\equiv$0.312 arcsec pixel$^{-1}$). The detector
was set to a binning of 2 $\times$ 2 in both the spatial and spectral
directions. Consequently, 512 increments, each 0\farcs624{} long gave
a projected slit length of 5\farcm32 on the sky. A TEK-1 CCD detector
was used for the 1998 data set with 1024$\times$1024 pixels$^2$ with
sides measuring 24 $\mu$m, using 2$\times$2 binning ($\equiv$0.62 arcsec
pixel$^{-1}$ and 3.5 \kms pixel$^{-1}$). For the 2001 and 2013 we used
a Marconi CCD detector with 2048 $\times$ 2048 square pixels, each 13.6
$\mu$m on a side. The detector was set to a binning of 2 $\times$ 2 in
both the spatial and spectra directions. Consequently, 1024
increments, each 0\farcs352{} long gave a projected slit length of
5\farcm47 on the sky.

We used a 90\,\AA\, and 50\,\AA\, bandwidth
filter to isolate the 87th and 114th orders containing the
\Ha\,$+$\,\NIIlam\, and \OIIIlam\, nebular emission lines,
respectively. We used a 70~$\mu$m{} ($\equiv$0.95\arcsec) slit, giving
a velocity resolution of 9.2 \kms{} ($\equiv$0.312 arcsec
pixel$^{-1}$ and 150 $\mu$m{} ($\equiv$1\farcs 9 and $\equiv$11.5
\kms) slit.

We obtained 23 consecutive and tightly spaced positions over the
NGC~6210. The slit positions are indicated and labeled in Figure~1 on
a WFPC~2 image of the NGC~6210 obtained from the HST archive. In order
to establish the exact position of the slit in each pointing, the slit
position on the sky was recorded with an automatic procedure available
in MES-SPM prior to the spectroscopic exposure.

The log of spectroscopy observation is given in Table 1 divided into
dates, number of spectra, exposure times, spectral range, resolution,
slit wide, position angle and name of the slit (see Figure~1). 
The high resolution data are available in San Pedro M\'artir Kinematic Catalogue
of Galactic Planetary Nebulae (http://kincatpn.astrosen.unam.mx)
\citep{Lopez:2012a}.

Due to saturation of several slit positions on the nebula, we obtained a 
new set of observation on August 13, 2015. We used the same detector as 
that used for the 2015 dataset, which was a E2V-4240 (Marconi 2)  detector 
with 2048$\times$2048 pixels$^2$ 
with sides measuring 13.5 $\mu$m, using 2$\times$2 binning 
($\equiv$0.531 arcsec pixel$^{-1}$. Spectra were obtained at 11 different 
positions across the nebula, with exposure times of 1800 seconds. The slit 
was oriented N-S.

An additional slit was placed outside the nebula 
to detect The outer halo of NGC 6210. The observation was carried out on 18 September 2019.  The spectrum was obtained in order 87 (H alpha and [N II] lines) with the 150 micron slit oriented at position angle of 56 degrees.  The detector was an E2V 2048x2048 CCD with 13.5 $\mu$/pix used in binning 3$\times$3.  The spectral dispersion was 0.084 \AA/pix and the plate scale 0.175\arcsec/pixel.

The characteristics of the observations are listed in Table 2.

\begin{table*}
\centering
\caption{Log  of Time-resolved  Observations of {\bf NGC~6210}.}
\begin{tabular}{l|ccccccc} \hline 
&  \multicolumn{7}{c}{Spectroscopy hight resolution/ 2.1m  OAN-SPM}            \\[0.1pt]
\hline 
Run    &   Exp. Time & Range  & slit & P. A.   & slit        \\
DD/MM/YYYY   &     frame         &    (s)   &  (\AA)    &    (\AA)  & (m$\mu$) & (degree) & name    \\[1pt]   \hline 
28/06/1998          & 30 & H$\alpha$  & 150 & 90 & o\\
28/06/1998          & 300 & H$\alpha$  & 150 & 90 & n\\
28/06/1998           & 1200 & H$\alpha$   & 150 & 90 & m,p,q \\
05/06/2003         & 1800 &  H$\alpha$   & 70 &0 & f,e,d,h,i,k,b\\ 
16/10/2003   &  1800 &   H$\alpha$    & 70 & $-$21 & t\\
16/10/2003    &  1800 &   H$\alpha$   & 70 & $-$68 & v\\
17/10/2003    &  1800 &   H$\alpha$  & 70 & 77 & w\\
13/06/2004   & 1800 &   O\sc{iii}  & 70 & $-$9 & r  \\  
13/06/2004   & 1800 &   O\sc{iii}  & 70 & $-$19 & s  \\  
14/06/2004    & 1800 &   S\sc{ii} &150 & $-$19 & s  \\
13/06/2004   & 1800 &   O\sc{iii}  & 70 & $-$56 & u  \\  
14/06/2004   & 1800 &   O\sc{iii}  & 150 & 0 & a,l  \\
21/05/2011     & 1800 & H$\alpha$  & 150 & 0 & g \\
21/05/2011     & 600 & O\sc{iii} & 150 & 0  & g\\
06/07/2013    &  1800 & H$\alpha$   & 150 & 0 & c,i,j \\
20/08/2005   & 3 &  1800 &   H$\alpha$, [OIII]    & 0.26 & 70 & 0 & h, j, k\\
\hline
\hline 
\end{tabular}
\label{table:pa5}
\end{table*}

\begin{table*}
\centering
\caption{Log  of Time-resolved  Observations of {\bf NGC~6210}. Epoch 2015}
\begin{tabular}{l|ccccccc} \hline 
&  \multicolumn{7}{c}{Spectroscopy hight resolution/ 2.1m  OAN-SPM}            \\[0.1pt]
\hline 
Run   &   Exp. Time & Range  & slit & P. A.   & slit        \\
DD/MM/YYYY   &     frame         &    (s)   &  (\AA)    &    (\AA)  & (m$\mu$) & (degree) & name    \\[1pt]   \hline 
18/08/2015  &  1800 &   H$\alpha$, [OIII]  & 70 & 0 & c, d, e, f, g\\
19/08/2015  &  1800 &   H$\alpha$, [OIII]   & 70 & 0 & b, a, i\\
19/08/2015  &  1800 &   H$\alpha$    & 70 & 0 & b, a, i\\
18/09/2019  &  1800 &   H$\alpha$    & 150 & 56 & \\

\hline
\hline 
\end{tabular}
\label{table:pa5}
\end{table*}

All data was reduced (bias removal 
and cosmic-ray removal) by using standard IRAF\footnote{IRAF is
  distributed by the National Optical Astronomy Observatory, which is
  operated by the Association of Universities for Research in
  Astronomy, Inc. under cooperative agreement with the National
  Science foundation}.
  This was followed by rectification and first-order wavelength 
  calibration of the two-dimensional spectra based on the comparison
  the spectrum
of a Th/Ar lamp to an accuracy of $\pm$1 \kms{} when converted to
radial velocity.  All spectra presented in this paper are corrected to
heliocentric velocity (\vhel). 
The Figure 3 shows the position-velocity ({\it P-V})  arrays of \Halam\,
\NIIlam \AA, \OIIIlam, respectively, where the heliocentric velocities are used,
and positions are specified as arcsecond offsets (x,y).

@@

\section{KINEMATICS}
\label{sec:kinematic}

The Figure 3 shows the \Halam{},  \OIIIlam{}, and \NIIlam{} velocity arrays, {\it P--V}, for all individual slit position, observed on 2015 epoch: \Ha{} {\it P--V} is shown on the left panel,  on
the middle panel is the corresponding \OIIIlam\, {\it P--V} and \NIIlam{} on the right panel. Spatial
offsets are in arcseconds.  The stellar continuum from the central star
has not been subtracted. We derive a heliocentric
systemic velocity, \vsys $=40$~\kms{} by using the slits position f (epoch 2003) and  which passes through the central star. The expansion velocity of the inner shell, v$_{exp}$ = 28 \kms.

@@

\begin{figure*}
\centering
  \includegraphics[width=0.7\textwidth]{tere-figs/Figure1}
  \caption{ }
\end{figure*}

\begin{figure*}
\centering
  \includegraphics[width=0.4\textwidth]{tere-figs/Figure2a}
  \caption{ }
\end{figure*}

@@

\begin{figure*}
\centering
\includegraphics[width=0.9\textwidth]{tere-figs/Figure8}
\includegraphics[width=0.42\textwidth]{tere-figs/N6210_ruso}
\quad
\includegraphics[width=0.427\textwidth]{tere-figs/opo9836f}
  \caption{}
\end{figure*}

\begin{figure*}
\centering
\includegraphics[width=0.9\textwidth]{tere-figs/Figure6}
  \caption{  }
\end{figure*}

\begin{figure*}
\centering
\includegraphics[width=0.9\textwidth]{tere-figs/rendija-c-cont}
  \caption{  }
\end{figure*}

\begin{figure*}
\centering

\includegraphics[width=0.4\textwidth]{tere-figs/slit_t_ha_nii}
\includegraphics[width=0.4\textwidth]{tere-figs/slit_v_ha_nii}
\includegraphics[width=0.4\textwidth]{tere-figs/slit_w_ha_nii}
\includegraphics[width=0.2\textwidth]{tere-figs/slit_u_oiii}
\includegraphics[width=0.2\textwidth]{tere-figs/slit_s_oiii}
\includegraphics[width=0.2\textwidth]{tere-figs/slit_r_oiii}

  \caption{  }
\end{figure*}

@@

>>>>

\usepackage{xcolor}
\usepackage{hyperref}
\usepackage{siunitx}                    ``
\usepackage{newtxtext}
\usepackage[stix2,smallerops]{newtxmath}
\usepackage{booktabs}
\hypersetup{colorlinks=True, linkcolor=blue!50!black, citecolor=black,
  urlcolor=blue!50!black}
\usepackage{etoolbox}
\robustify\bfseries
\robustify\itshape

@@

@@

\begin{abstract}
  A comprehensive kinematic study has been carried out on the planetary nebula (PN) NGC 6210. Multiple long-slits, echelle spectra have been obtained over the face of this nebula mapping its full shell structure, the opposite pairs of extended, twisted arms and the bipolar collimated outflows and bullets. Public {\it HST} imagery has been used to identify kinematic elements with structural components. The long-slit spectroscopic information has been combined into channel maps that greatly facilitate visualizing the otherwise intricate expansion pattern of this  planetary nebula. The global morphology of NGC 6210 is  reminiscent of other planetary nebulae with extended X-ray emission, suggesting the presence of a central hot bubble produced by shocked stellar wind. In spite of the dramatic structure of this PN substantial expanding radial motions are only found in the material surrounding the central, inner shell. The relatively slow radial motions of the collimated outflows and the bullet-like knots indicate that they are  moving away from the core very close to the plane of the sky, nearly perpendicular to the line of sight. The direction and ages of the various components, derived from their expanding velocities indicate that NGC 6210 has been strongly influenced by the  action of a bipolar, rotating, episodic outflow.
\end{abstract}

@@

\section{INTRODUCTION}
\label{sec:introduction}
NGC 6210 is a bright and relatively large PN in the northern sky. The first photographic image of NGC 6210 was published by \citet{Duncan:1937a}. In that old, fuzzy image a pair of twisted arms protruding in opposite directions from a blurred,  bright, nebular core are apparent. Several studies have discussed the inner kinematic motions of this object, e.g. \citet{Osterbrock:1966a, Weedman:1968a, Becker:1984a, Icke:1989a}. However, none of those studies had the spatial coverage and spectral resolution needed to produce a reliable spatio-kinematic model of this PN. 
The best kinematic study to date on NGC 6210 is probably the one published by \citet{Phillips:1996a}  who obtained 11 long-slits centered on the nebular core, with each slit at different position angles thus providing a good spatial coverage but with only limited, low and medium spectral resolution. From these data they suggested a reasonable, though somewhat basic, kinematic model for the ionized nebula. NGC 6210 jumped into fame thanks to the {\it HST} images obtained during the 1996 - 1998 period under the observing programs 6347 (K. Borkowski), 6792 (R. Rubin), 7501 (A. Hajian)  and 11122 (B. Balick). These images showed for the first time the extraordinarily complex morphology of NGC 6210. Incidentally the  {\it HST} and ground-based imagery prompted the nickname "the turttle" for this nebula, given the shape of its main body and extended arms that resemble the shell of a turttle with a long neck and its fins. The distance estimates to NGC 6210 vary roughly from 1.5 to 2.0 kpc. \citep{Hajian:1995a} derive a distance to NGC 6210 from VLA 6 cm  expansion parallax measurements of 1.57$\pm0.40$ kpc, considering an expansion velocity of 23 \kms, whereas Frew (20XX) derives a distance of kpc to this object from its H$\alpha$ surface brightness method.

In this work we present a thorough spectral mapping of all the morphological elements of the PN obtained at high spectral resolution that allow 
disentangling the various spatio-kinematic components and provide an overall view of its structure and evolution. Radial velocities are combined with 2-epoch HST \oiii ~and~ \nii ~archive image sets (http://archive.stsci.edu/) to derive proper motions. Individual long-slit echelle spectra obtained over 5 observing runs are analyzed as position - velocity data and combined into emission and velocity channel maps that serve as an excellent tool to visualize and understand the peculiar velocity fields, of the several structural components.

In Section 2, we describe our observations and the data-reduction steps, In Section 3 we describe the proper motions, In Section 4, we discuss the kinematic components from slit spectra. In next section, we present the dynamical ages and finally in section 5 we present the discussion and conclusion.

@@

Long-slit, echelle, spectroscopic observations of the nebula NGC~6210
were performed with the 2.1~m telescope at the Observatorio
Astron\'omico Nacional at San Pedro M\'artir, (OAN-SPM), Baja
California, M\'exico. We used the Manchester Echelle Spectrometer
(MES-SPM) \citep{Meaburn:2003a} on the 2.1 m telescope in its $f$/7.5
configuration.  The MES-SPM is a long-slit, echelle spectrometer that
has no cross-disperser; it isolates single orders using interference
filters. We used a 90\,\AA\, and 50\,\AA\, bandwidth
filter to isolate the 87th and 114th orders containing the
\Ha\,$+$\,\NIIlam\, and \OIIIlam\, nebular emission lines,
respectively. Spectroscopic data on NGC 6210 was collected within our regular observing programs over nine observing runs from 1998 till 2019. Table 1 shows the log of observations. In the first column the dates for the individual observing runs are listed; the second column lists the exposure times mostly used in the corresponding run; the third column lists the spectral range of the observations, namely, $H\alpha$ + [N II] and/or [O III]; the fourth column lists the slit width in microns and the fifth column the position angle of the slit. Finally, in the sixth column the nomenclature used to identify the slits obtained in each run is indicated. In the end 34 slit positions from all the runs were selected to analyse the kinematic structure of NGC 6210. The location of these slits are presented and labeled against an [N II] $\rm HST$ archive image in Figure 1. In the top panel the vertical solid lines represent the set of 11 slits obtained in the 2015 run. The vertical dashed lines represent 12 slit positions from other runs, in some cases they are highly coincident in location with the 2015 set but selected for reasons such as different exposure times or spectral range. The middle panel shows 5 slit positions oriented E - W and the bottom panel 6 slit positions obtained at an angle intended to record specific features, such as collimated outflows or the twisted arms of the nebula. In order to establish the exact position of the slit in each pointing, the slit position on the sky was recorded with an automatic procedure available in MES-SPM prior to the spectroscopic exposure.

During the 11 years span of the observations different detectors were used with MES. Their general characteristics are described below

A TEK-1 CCD detector was used for the 1998 data set with 1024 $\times$ 1024 square pixels, each 24$\mu$m on a side, $\equiv$0.312 arcsec pixel$^{-1}$. A 2$\times$2 binning was set, consequently, 512 increments, each 0\farcs624{} long gave a projected slit length of 5\farcm32 on the sky. For slits 150 $\mu$m ($\equiv$1.9\arcsec) and 70 $\mu$m ($\equiv$0.95\arcsec) wide the spectral resolution is 11.5 \kms and 4.6 \kms, respectively).

For the 2003, 2004 and 2011 runs of the present observations, we used a SITE-3 CCD detector. That detector had the same dimensions (1024 $\times$ 1024 square pixels, each 24$\mu$m on a side) as the Tek-1.

For the 2011 and 2013 runs we used a Marconi CCD detector with 2048 $\times$ 2048 square pixels, each 13.5
$\mu$m on a side. The detector was set to a binning of 2 $\times$ 2 in both the spatial and spectra directions. Consequently, 1024
increments, each 0\farcs352{} long gave a projected slit length of 5\farcm47 on the sky. For a 70~$\mu$m{} a velocity resolution of 5.9 \kms{} is obtained in this case, whereas the 150~$\mu$m slit yields 11.9 \kms{}.

During data reduction we noticed saturation in several slit positions near the core of the nebula. Therefore we decided to obtain in 2015 a new set of uniform and evenly spaced observation. \Ha{}, \NIIlam{} and \OIIIlam{} where observed for each slit position. We used the Marconi 2 (E2V-4240) detector with an identical set up as the one used in the 2013 run. Long-slit spectra were obtained at 11 different positions across the nebula, stepped from east to west and the slit oriented N -S.

Figure 2 shows the position-velocity ({\it P-V})  arrays of \Halam\,
\NIIlam \AA, and \OIIIlam, respectively, for the 11 individual positions from the 2015 run. \Ha{} {\it P--V} is shown on the left panel,  on
the middle panel is the corresponding \OIIIlam\, {\it P--V} and \NIIlam{} on the right panel. The velocity scale is heliocentric 
and the scale along the slit is specified as arcsecond offsets from the core position. We derive a heliocentric
systemic velocity, \vsys $=40$~\kms{} by using the slits position f (epoch 2003) that passes through the central star and is not saturated.

A final slit was placed on the northern outskirts of the nebula 
in the outer halo of NGC 6210 where filaments had been clearly detected in some of our deep images. The observation was carried out on 18 September 2019.  The spectrum was obtained in order 87 ($H\alpha$ and [N II] lines) with the 150 micron slit oriented at a position angle of 56 degrees.  The detector was again the Marconi 2048x2048 with a 3$\times$3 binning.  The spectral dispersion was 0.087 \AA/pix and 0.525\arcsec/pixel.

Figure 3 shows the location of the slit in the halo of NGC 6210  against a deep ground-based image. The bottom panel shows the large scale view with the location of the slit indicated. The top right panel is an enhanced contrast image of the region that intersects the slit, the regions shows a knotty, curved structure. The top left panel shows the corresponding spectrum from the 2019 run.

Most of the spectra presented here are included in the San Pedro M\'artir Kinematic Catalogue of Galactic Planetary Nebulae and are publicly available as pdf or fits files (http://kincatpn.astrosen.unam.mx)
\citep{Lopez:2012a}.

 \begin{figure}
\centering
  \includegraphics[width=0.4\textwidth]{tere-figs/Figure-i+s}
  \caption{ }
\end{figure}

\begin{table}
\centering
%\caption{Log  of Time-resolved  Observations of {\bf NGC~6210}.}
\begin{tabular}{l|cccccccc} \hline 
&  \multicolumn{7}{c}{Spectroscopy hight resolution/ 2.1m  OAN-SPM}            \\[0.1pt]
\hline 
Run    &   Exp. & Range  & Slit & P. A.   & slit        \\
Epoch   &     Time        &    (s)   &  (\AA)    &    (\AA)  & (m$\mu$)     \\[1pt]   \hline 
28/06/1998          & 30 & H$\alpha$  & 150 & 90 & O\\
28/06/1998          & 300 & H$\alpha$  & 150 & 90 & N\\
28/06/1998           & 1200 & H$\alpha$   & 150 & 90 & M,P,Q \\
05/06/2003         & 1800 &  H$\alpha$   & 70 &0 &     B,D,E,F,H,I,K\\ 
16/10/2003   &  1800 &   H$\alpha$    & 70 & $-$21 & T\\
16/10/2003    &  1800 &   H$\alpha$   & 70 & $-$68 & V\\
17/10/2003    &  1800 &   H$\alpha$  & 70 & 77 & W\\
13/06/2004   & 1800 &   O\sc{iii}  & 70 & $-$9 & R  \\  
13/06/2004   & 1800 &   O\sc{iii}  & 70 & $-$19 & S  \\  
14/06/2004    & 1800 &   S\sc{ii} &150 & $-$19 & S  \\
13/06/2004   & 1800 &   O\sc{iii}  & 70 & $-$56 & U  \\  
14/06/2004   & 1800 &   O\sc{iii}  & 150 & 0 & A,L  \\
21/05/2011     & 1800 & H$\alpha$  & 150 & 0 & G \\
21/05/2011     & 600 & O\sc{iii} & 150 & 0  & G \\
06/07/2013    &  1800 & H$\alpha$   & 150 & 0 & C,I,J \\
18/08/2015  &  1800 &   H$\alpha$, [OIII]  & 70 & 0 & C,D,E,F,G\\
19/08/2015  &  1800 &   H$\alpha$, [OIII]   & 70 & 0 & B,A,I\\
20/08/2015   &  1800 &   H$\alpha$, [OIII] & 70 & 0 & H,J,K\\
18/09/2019  &  1800 &   H$\alpha$    & 150 & 56 & \\

\hline
\end{tabular}
\label{table:pa5}
\end{table}

All data was reduced  by using standard IRAF\footnote{IRAF is
  distributed by the National Optical Astronomy Observatory, which is
  operated by the Association of Universities for Research in
  Astronomy, Inc. under cooperative agreement with the National
  Science foundation} routines. Bias and cosmic-ray removal was followed by rectification and first-order wavelength 
  calibration of the two-dimensional spectra based on the comparison
  the spectrum
of a Th/Ar lamp to an accuracy of $\pm$1 \kms{} when converted to
radial velocity.  All spectra presented in this paper are corrected to
heliocentric velocity (\vhel).

The morphology of NGC 6210 varies significantly in high and low excitation images. In Figure 4 we show in the lower panels [O III] (left) and [N II] (right) archive $\rm HST$ images that clearly reveal the different structures from each ion.
The top panels are the corresponding images processed with a natural logarithm and edge detection algorithm where the most salient features are indicated and labeled for reference in the next sections.

@@

%\section{KINEMATICS}
%\label{sec:kinematic}

%The Figure 3 shows the \Halam{},  \OIIIlam{}, and \NIIlam{} velocity arrays, {\it P--V}, for all individual slit position, observed on 2015 epoch:   The stellar continuum from the central star
%has not been subtracted. . The expansion velocity of the inner shell, v$_{exp}$ = 28 \kms.

@@

%\begin{figure*}
%\centering
%  \includegraphics[width=0.7\textwidth]{tere-figs/Figure1}
%  \caption{ }
%\end{figure*}

@@

\begin{figure*}
\centering
\includegraphics[width=0.4\textwidth]{tere-figs/Figure6}
  \caption{  }
\end{figure*}

\begin{figure}
\centering
%\includegraphics[width=0.7\textwidth]{tere-figs/Figure8}
%\includegraphics[width=0.327\textwidth]{tere-figs/opo9836f}
%\includegraphics[width=0.32\textwidth]{tere-figs/N6210_ruso}
\includegraphics[width=0.6\textwidth]{sketch}

  \caption{}
\end{figure}

%\begin{figure*}
%\centering
%\includegraphics[width=0.9\textwidth]{tere-figs/rendija-c-cont}
%  \caption{  }
%\end{figure*}

%\begin{figure*}
%\centering

%\includegraphics[width=0.4\textwidth]{tere-figs/slit_t_ha_nii}
%\includegraphics[width=0.4\textwidth]{tere-figs/slit_v_ha_nii}
%\includegraphics[width=0.4\textwidth]{tere-figs/slit_w_ha_nii}
%\includegraphics[width=0.2\textwidth]{tere-figs/slit_u_oiii}
%\includegraphics[width=0.2\textwidth]{tere-figs/slit_s_oiii}
%\includegraphics[width=0.2\textwidth]{tere-figs/slit_r_oiii}

  %\caption{  }
%\end{figure*}

@@

Previously, \citet{Schonberner:2018a} have also studied the proper motions
in the Turtle Nebula, based on the same HST observations as those used here.
They employ a magnification method, which implicitly assumes a homologous expansion,
finding a magnification factor over the 10 to 11-year baseline
of \(M = 1.0085 \pm 0.0005\) for the inner high-ionization shells
(which they call the cavity rim)
and \(M = 1.0055 \pm 0.0005\) for the low-ionization knot complexes (see their Table~2).
After conversion to velocity units (assuming a distance of \SI{2}{kpc}),
these are consistent within the stated errors with our own average results for each of these two nebular components.
